### 2017-10-22 paper/sections/0_abstract.tex
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

HPC developers are abandoning POSIX because the synchronization and
serialization overheads of providing strong consistency and durability are too
costly -- and often unnecessary -- for their applications.  Unfortunately,
designing near-POSIX file systems excludes applications that rely on strong
consistency or durability, forcing developers to re-write their applications or
deploy them on a different system.  We present a file system and API that lets
clients specify their consistency/durability requirements and assign them to
subtrees in the namespace, allowing users to optimize subtrees within the same
namespace for different workloads.  We draw conclusions about the performance
impact of unexplored consistency/durability metadata designs and show that
maintaining strong consistency can cause about a 100\(\times\) slow down
compared to relaxed consistency and no durability. Comparatively, merging
updates after a period of relaxed consistency (less than a 10\(\times\) slow
down) and maintaining durability (about a 10\(\times\) slow down) have more
reasonable costs.

>>>>

Data center scale application developers are abandoning POSIX because the file
system metadata synchronization and serialization overheads of providing strong
consistency and durability are too costly -- and often unnecessary -- for their
applications.  Unfortunately, designing near-POSIX file systems excludes
applications that rely on strong consistency or durability, forcing developers
to re-write their applications or deploy them on a different system. Mounting
multiple systems in the global namespace forces users to (1) provision separate
storage clusters and (2) manually move data across systems.  We present a
framework and API that lets clients dynamically specify their
consistency/durability requirements and assign them to subtrees in the
namespace, allowing users to optimize subtrees within the same namespace for
different workloads.  We confirm the performance benefits of techniques
presented in related work but also explore new consistency/durability metadata
designs, all over the same storage system.  By custom fitting a subtree to a
create-heavy application, we show an 8\(\times\) speedup and scale further
(twice as many servers) than the system we implemented our prototype on.

### 2017-10-22 paper/sections/1_introduction.tex
paper: intro to the evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}

\caption{For performance, clients can relax consistency on their subtree
({e.g.}, HDFS) or even decouple the subtree and move it locally ({e.g.},
BatchFS, RAMDisk). Decoupled subtrees can further relax durability semantics
for even better performance. Clients that require stronger guarantees ({\it
e.g.,} POSIX) can still reside in the same namespace.
}\label{fig:subtree-policies}

\end{figure}

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many batch style jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
patterns).

@@

% What did we do
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a setup in
our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art HPC
architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
paper, although the HDFS semantics and their performance is explored
in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
consistency because it  does not guarantee that all updates are not immediately
seen by all servers; the BatchFS and RAMDisk subtrees are decoupled from the
global namespace and have similar consistency/durability behavior to those
systems; and the POSIX subtree retains the rigidity of POSIX's strong
consistency.

@@

  \item a prototype that lets developers choose from a range of
  consistency and durability semantics (9 permutations), allowing them to dynamically custom
  fit the storage system to the application.

  \item an API for selecting consistency/durability policies and assigning
  them to subtrees in the file system namespace.

  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
  previously unexplored metadata designs, all implemented using Cudele.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.  Our
prototype is built on Ceph because it was originally designed for HPC and has
proven to be useful for HPC workloads~\cite{}. Several features were motivated
by HPC requirements~\cite{weil:osdi2006-ceph}, so the lessons are useful (and
can be implemented) in HPC environments.

>>>>

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many batch style jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N and N:1 create
patterns~\cite{bent_plfs_2009}).

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{ Illustration of subtrees with different semantics co-existing in a
global namespace.  For performance, clients can relax consistency on their
subtree (HDFS) or even decouple the subtree and move it locally (BatchFS,
RAMDisk). Decoupled subtrees can further relax durability for even better
performance.  Clients that require stronger guarantees (POSIX) can still reside
in the same namespace.  }\label{fig:subtree-policies}
\end{figure}

@@

% What did we do
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a potential
setup in our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art
architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
paper, although the HDFS semantics and their performance is explored
in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
consistency because it does not guarantee that all updates are immediately seen
by all servers; the BatchFS and RAMDisk subtrees are decoupled from the global
namespace and have similar consistency/durability behavior to those systems;
and the POSIX subtree retains the rigidity of POSIX's strong consistency.
Subtrees without policies inherit the consistency/durability semantics of the
parent and future work will examine embeddable or composable policies.

@@

  \item A framework and API for assigning consistency/durability policies 
  to subtrees in the file system namespace; this lets users dynamically navigate
  the tradeoffs of different metadata protocols.

  \item This framework lets subtrees with different semantics co-exist in a
  global namespace. This is different than multiple mount points because
  there is no need to provision storage clusters for applications
  ({\it e.g.}, HDFS, CephFS) or move data between these systems.

  \item A prototype that lets developers custom fit subtrees to applications
  dynamically; this helps data center systems scale further than systems that
  use one strategy for the entire namespace.

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). In our prototype, we get an 8\(\times\) speedup and
can scale to twice as many servers when we assign a more relaxed form of
consistency and durability to a subtree with a create heavy workload.
Meanwhile, other subtrees in the namespace maintain strong consistency and
durability with RPCs per operation.

We build on Ceph because it is used in cloud-based and data center
systems and has a presence in HPC~\cite{}.  More importantly,  we use Ceph as a
platform for exploring related work over the same storage system, which
facilitates an apples-to-apples comparison of the strategies themselves.

### 2017-10-22 paper/sections/2_related_work.tex
paper: draft 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% General
The bottlenecks associated with accessing POSIX file system metadata are not limited
to HPC workloads and the same challenges that plagued these systems for years are
finding their way into the cloud. Workloads that deal with many small files
({\it e.g.}, log processing and database
queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
simultaneous clients ({\it e.g.}, MapReduce
jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
the metadata service. The biggest challenge is that whenever a file
is touched the client must access the file's metadata and maintaining a file
system namespace imposes small, frequent accesses on the underlying storage
system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
system metadata is a well-known problem and solutions for scaling data IO do
not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
abad:techreport2012-fstrace, abad:ucc2012-mimesis,
alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
for improving the performance of metadata access.

\subsection{Metadata Load Balancing}

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incur extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to maneuver
tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requires
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to directories in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoritative metadata server responds.  These protocols are discussed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

%For metadata writes it does a distributed transaction; monotonic writes with
%concurrent clients fail and do pessimistic locking through a mediated lock
%server to ensure strong consistency, non-monotonic writes grab locks at every
%server. Zooming in on monotonic writes: if permissions increase it executes on
%a primary then non-primary, if permissions decrease it executes on all
%non-primary then on primary.

The conclusion we have drawn from this related work is that metadata protocols
have a bigger impact on performance and scalability than load balancing.  
Understanding these protocols helps load balancing and gives us a better
understanding of the metrics we should use to make migration decisions ({\it
e.g.}, which operations reflect the state of the system), what types of
requests cause the most load, and how an overloaded system reacts ({\it e.g.},
increasing latencies, lower throughput, etc.).

\subsection{Relaxing POSIX}

>>>>

% General
The bottlenecks associated with accessing POSIX file system metadata are not limited
to HPC workloads and the same challenges that plagued these systems for years are
finding their way into the cloud. Workloads that deal with many small files
({\it e.g.}, log processing and database
queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
simultaneous clients ({\it e.g.}, MapReduce
jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
the metadata service. The biggest challenge is that whenever a file
is touched the client must access the file's metadata and maintaining a file
system namespace imposes small, frequent accesses on the underlying storage
system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
system metadata is a well-known problem and solutions for scaling data IO do
not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
abad:techreport2012-fstrace, abad:ucc2012-mimesis,
alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. 
%There are two approaches
%for improving the performance of metadata access.
%
%\subsection{Metadata Load Balancing}
%
%% approaches to load balancing (strong = more resuorces per unit work, weak =
%% fixed resource per work unit)
%% CITEME
%One approach for improving metadata performance and scalability is to alleviate
%overloaded servers by load balancing metadata IO across a cluster. Common
%techniques include partitioning metadata when there are many writes and
%replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
%directories and clients write to different partitions by grabbing leases and
%caching ancestor metadata for path traversal; it does well for strong scaling
%because servers can keep more inodes in the cache which results in less RPCs.
%Alternatively, ShardFS replicates directory state so servers do not need to
%contact peers for path traversal; it does well for read workloads because all
%file operations only require 1 RPC and for weak scaling because requests will
%never incur extra RPCs due to a full cache.  CephFS employs both techniques to
%a lesser extent; directories can be replicated or sharded but the caching and
%replication policies do not change depending on the balancing technique.
%Despite the performance benefits these techniques add complexity and jeopardize
%the robustness and performance characteristics of the metadata service because
%the systems now need (1) policies to guide the migration decisions and (2)
%mechanisms to address inconsistent states across servers.
%
%% policies
%Setting policies for migrations is arguably more difficult than adding the
%migration mechanisms themselves.  For example, IndexFS and CephFS use the
%GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
%threshold and using lazy synchronization to redirect queries to the server that
%``owns" the targeted metadata.  Determining when to partition directories and
%when to migrate the directory fragments are policies that vary between systems:
%GIGA+ partitions directories when the size reaches a certain number of files
%and migrates directory fragments immediately; CephFS partitions directories
%when they reach a threshold size or when the write temperature reaches a
%certain value and migrates directory fragments when the hosting server has more
%load than the other servers in the metadata cluster. Another policy is when and
%how to replicate directory state; ShardFS replicates immediately and
%pessimistically while CephFS replicates only when the read temperature reaches
%a threshold.  There is a wide range of policies and it is difficult to maneuver
%tunables and hard-coded design decisions.
%
%% addressing inconsistency
%In addition to the policies, distributing metadata across a cluster requires
%distributed transactions and cache coherence protocols to ensure strong
%consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
%directory state and uses optimistic concurrency control for conflicts; namely
%it does the operation and if there is a conflict at verification time it falls
%back to two-phase locking.  Another example is IndexFS's inode cache which
%reduces RPCs by caching ancestor paths -- the locality of this cache can be
%thrashed by random reads but performs well for metadata writes. For
%consistency, writes to directories in IndexFS block until the lease expires
%while writes to directories in ShardFS are slow for everyone as it either
%requires serialization or locking with many servers; reads in IndexFS are
%subject to cache locality while reads in ShardFS always resolve to 1 RPC.
%Another example of the overheads of addressing inconsistency is how CephFS
%maintains client sessions and inode caches for capabilities (which in turn make
%metadata access faster). When metadata is exchanged between metadata servers
%these sessions/caches must be flushed and new statistics exchanged with a
%scatter-gather process; this halts updates on the directories and blocks until
%the authoritative metadata server responds.  These protocols are discussed in
%more detail in the next section but their inclusion here is a testament to the
%complexity of migrating metadata.
%
%%For metadata writes it does a distributed transaction; monotonic writes with
%%concurrent clients fail and do pessimistic locking through a mediated lock
%%server to ensure strong consistency, non-monotonic writes grab locks at every
%%server. Zooming in on monotonic writes: if permissions increase it executes on
%%a primary then non-primary, if permissions decrease it executes on all
%%non-primary then on primary.
%
%The conclusion we have drawn from this related work is that metadata protocols
%have a bigger impact on performance and scalability than load balancing.  
%Understanding these protocols helps load balancing and gives us a better
%understanding of the metrics we should use to make migration decisions ({\it
%e.g.}, which operations reflect the state of the system), what types of
%requests cause the most load, and how an overloaded system reacts ({\it e.g.},
%increasing latencies, lower throughput, etc.).

%\subsection{Relaxing POSIX}

### 2017-10-22 paper/sections/6_results.tex
paper: add scalability block/allow
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: touch intro and abstraction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
client. The size of the client's journal scales with the number of
updates.\label{fig:behavior-journal-size}}
\end{figure}

@@

{\it Journal Size:} Figure~\ref{fig:behavior-journal-size} shows the
amount of storage per journal update (\(y\) axis) for the range of file creates
we tested (\(x\) axis). The increase in file size is linear with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
payloads of this size on an HPC network are reasonable.\\

@@

\noindent\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building evaluating different
consistency/durability semantics.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
\caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and
``block" API, users can isolate directories from interfering clients. Variance
with blocking turned on is the same as ``isolated" from
Figure~\ref{fig:overhead-b}.
\label{fig:slowdown-allow-block}}

>>>>

%\begin{figure}[tb]
%\centering
%\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
%\caption{ [\href{https://...}{source}] Microbenchmark: storage footprint on
%client. The size of the client's journal scales with the number of
%updates.\label{fig:behavior-journal-size}}
%\end{figure}

@@

{\it Journal Size:} We measure the amount of storage per journal update to be
about 2.5\text{KB}. The storage footprint scales linearly with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
payloads of this size in most HPC/data cente networks are reasonable.\\

@@

\noindent\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
\caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and
``block" API, users can isolate directories from interfering clients. Variance
with blocking turned on is the same as ``isolated" from
Figure~\ref{fig:overhead-b}.
\label{fig:slowdown-allow-block}}
